% https://www.monash.edu/graduate-research/examination/publication
% https://www.monash.edu/rlo/graduate-research-writing/write-the-thesis
% https://www.monash.edu/rlo/graduate-research-writing/write-the-thesis/writing-the-thesis-chapters/structuring-a-long-text
\ack{
\addtocontents{toc}{}  % Add a gap in the Contents, for aesthetics


This work would not have been possible without the support and guidance of Rory and Eric. Thank you so much for everything you've done to help me become an independent researcher!

Rory, while thinking back on my Astro career, I would be negligent if I didn't start in the summer of 2016, when you invited me to join the LIGO Surf. I had a wonderful time working with you that summer, and I hoped that we could team up again in the future. Fortunately, you agreed to advise me for my PhD a few years later! So, thank you, Rory, for taking a gamble on me all those years ago, and for your continued encouragement and direction ever since! You’re an incredible mentor.

Eric, working with you has been an absolute pleasure! You have been extremely supportive and patient with me, while also challenging me to improve as a scientist. Your meticulous paper edits, your random anecdotes sprinkled into research meetings, and your calm demeanour whilst I frantically ranted about a new bug have really helped me get through this PhD. 

Both of you have been incredible supervisors! 

To Reinhold, Hans and Mike, you three have been the most supportive friends I could have asked for during these past three years (I made sure to use a legible font with dark ink for your sake, Reinhold). Thanks for making my time in Australia so awesome! Im looking forward to our next hiking trip so I can stick some more wasabi on Mike’s Oreos and yell ``almost there'' when we’re really not. 

That said, I'd like to express my gratitude to my parents for helping me grow into the person I am today, and to Didi, who has been the most incredible sister. Things were hard being so far away and isolated during the pandemic, but your support and encouragement helped me get through it!

And Rebecca – you have been truly extraordinary throughout my PhD, and I’m so grateful to you for helping me through this. I’ll defer further, soppier gratitude to a more direct mode of communication. 

Finally, I would like to thank Shaun T and Nikolai Belinski, who were instrumental in getting me through the 262 days of lockdown. Though you both could be annoying sometimes, you helped make the lockdown more bearable. On that note, I think it would be perfectly happy if I never heard Shaun T yell ``time for power pushuuuups!'' again. 

This research was supported by an Australian Government Research Training Program (RTP) Scholarship and OzGrav, the ARC centre of excellence for gravitational-wave discovery.
}
