% https://www.monash.edu/graduate-research/examination/publication
% https://www.monash.edu/rlo/graduate-research-writing/write-the-thesis
% https://www.monash.edu/rlo/graduate-research-writing/write-the-thesis/writing-the-thesis-chapters/structuring-a-long-text
%https://www.monash.edu/graduate-research/examination/publication
\Declarationpublication{
\addtocontents{toc}{}  % Add a gap in the Contents, for aesthetics

I, \authors, hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the
thesis.

This thesis includes three original papsers published in peer reviewed journals and two submitted publications.
These papers are provided in Chapters~\ref{ch.bcr}-\ref{ch.tess}.
The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the
 under the  supervision of Dr Rory Smith and Professor Eric Thrane.

Although I was responsible for some results, and software development of the paper that constitutes \cref{ch.pbilby}, I do not claim to be the principal contributor to this work. Due to the collaborative nature of this project, it is not examinable under the university rules, however, it is included to preserve the narrative of the thesis. I, the student, working within the School of Physics and Astronomy under the supervision of Dr Rory Smith and Professor Eric Thrane, played a majority role in contributing to the papers constituting Chapters~\ref{ch.bcr}-\ref{ch.tess}. My contributions to the papers presented in Chapters 2–7 are the following:


The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into
team-based research.

In the case of \emph{(insert chapter numbers}) my  contribution to the work involved the following:


Here should be the table!

I have / have not renumbered sections of submitted or published papers
in order to generate a consistent presentation within the thesis.

\textbf{Student name:} Avi Vajpeyi
\rule[1em]{25em}{0.5pt}

\textbf{Student signature:}
\rule[1em]{25em}{0.5pt}

\textbf{Date:}
\rule[1em]{25em}{0.5pt}

I, Rory Smith, hereby certify that the above declaration correctly reflects the
nature and extent of the student's and co-authors' contributions to this
work.
In instances where I am not the responsible author I have
consulted with the responsible author to agree on the respective
contributions of the authors.

\textbf{Primary Supervisor name:} Rory Smith
\rule[1em]{25em}{0.5pt}  % This prints a line for the signature

\textbf{Primary Supervisor signature:}
\rule[1em]{25em}{0.5pt}  % This prints a line for the signature

\textbf{Date:}
\rule[1em]{25em}{0.5pt}  % This prints a line for the signature















}
